---
title: "esm206-lab6-ac"
author: "Annie Combs"
date: "November 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test

We'll make our own samples using a pseudorandom generator

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

```{r}
hist(gp_1)

hist(gp_2)
```

Try a t test
```{r}
t.test(gp_1, gp_2)
```
The p value means that there is a 19.8% chance that the sample means are at least this different by random chance, assuming that these two samples come from populations with the same mean.

We will fail to reject (retain) the null hypothesis. >> there is no significant difference in means between group 1 and group 2 (say that on final)

Warning: people get weirdly upset if you say "accept the null". If accept or reject the null comes up in your final, you will LOSE POINTS. 

Compare this outcome to a rank-based test.

## Mann Whitney U unpaired rank based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```
p value means that there is a 28% chance that your sample ranks will be at least this different assuming that your samples came from populations with the same rank.
