---
title: "esm206-lab6-ac"
author: "Annie Combs"
date: "November 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test

We'll make our own samples using a pseudorandom generator

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

```{r}
hist(gp_1)

hist(gp_2)
```

Try a t test
```{r}
t.test(gp_1, gp_2)
```
The p value means that there is a 19.8% chance that the sample means are at least this different by random chance, assuming that these two samples come from populations with the same mean.

We will fail to reject (retain) the null hypothesis. >> there is no significant difference in means between group 1 and group 2 (say that on final)

Warning: people get weirdly upset if you say "accept the null". If accept or reject the null comes up in your final, you will LOSE POINTS. 

Compare this outcome to a rank-based test.

## Mann Whitney U unpaired rank based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
# ?wilcox.test in console to check values
# mwu$p.value (or other value) into console to pull individually
```
p value means that there is a 28% chance that your sample ranks will be at least this different assuming that your samples came from populations with the same ranks (median).

No significant difference in ranks/medians between group 1 and group 2.

Median Scores for group 1: (M = 14) and group 2 (M = 12) did not differ significantly (Mann Whitney U test: U(df) = 86, p = 0.28).

## Linear Regression

Simple linear regression (single dependent variable and a single independent variable)

```{r, include = FALSE} 
# Make an exploratory plot of penguin body mass vs flipper length (x axis)

ggplot(data = penguins, aes( x = flipper_length_mm, y = body_mass_g))+
  geom_point()+
  theme_minimal()
```

